MLH1 Results Outline

### 2020-02-23 doc/Results.rmd
must make progress

total SC t test
 p = `r Q2tot.sc_t.test_High.Low$p.value`
**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

@@

within subspecies

musc-short

@@

mol-long

@@

across subspecies:

dom - musc

dom-mol

<this might indicate -there's error in the total SC algorithm
- mouse effect
- low power / sampling bias in reduced data set
- total SC is capturing differences in the XY signal which is excluded from the bivalent data set>

@@

```{r Q2.totSC.glm, echo=FALSE}

#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel

@@

#sum.Q2.glm_mus.av_totSC_M1.coef <-

@@

glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)

@@

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)

#summary(glm_SC.male_Musc_long.biv)

@@

<we ran models to start examine strain specific effects ... modeling the fixed effects of subspeices and strain.>

Thse are our prediction for the glm models -- based on the variation of the gwRR's

@@

(the subspecies effect is tested in M1)
For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)

strain effects from M1; G, LEW, MSM, SKIVE and TOM

@@

<Compare by chrm class, but only for strains with similar proportions>

We purpose we think that running the models for comparing SC means by chromosome class and divided by rec group. (since the chromosome proportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )

The null hypothesis is that these won't be highly significant strain effects.
If there are still strain effects for these models, that would be evidence that there has been evolution for SC length compaction across these strains

@@

The CZECH and AST are higher than KAZ

The number of cells and mice is difference across the low group strains, 
-- permute the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)

above are glm tables for mouse.av SC lengths for each strain.

Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.

@@

<!--
blocks of other male Q2 - SC lengths

what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.

-->

> **Summary**

When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression. I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length. (IE the scatter plots of SC length, binned by chrm class on the X).

@@

When all male mice are used, the predictive power is greater, than when just the Musc strains are used. When, just the Musc strain are used, The mouse mean SC is slightly significant in predicting if a mouse is in the high or low (should I consider running on female too?)

> Is the prediction, high rec musc male strains have long SC met?

In a logistic regression, mouse average SC length is slightly predictive telling if a mouse is in a high or low Rec strain.  I couldn't get the Mixed models working for the male polymorphism predictions...

@@

These are the plots of the logistic regression.

@@

the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group

@@

```{r Q2.SC.Length.stuff, echo=FALSE, include=FALSE}
#REMOVE THIS, COMPARES FULL BIVDATA ACROSS CHRM CLASS

#all pooled SC lengths
#WSB     G   LEW  MOLF   KAZ CZECH 
#    5     9     5     4     3     3

#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14

#G sample 5 mice

#break down of mice in the low rec dataset

#glm.1CO
Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)

#glm.2CO
Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
#fewer significant differences

Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
#all but KAZ are slightly significant, G and CZECH more so

#glm.3CO
Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
#not sig differences
summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)

#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)

#all pooled SC lengths
lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE) 
lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)
Q1lmer.MM_SC_short_results
```

```{r Q2.SC.ploy.male, echo=FALSE, include=FALSE}
#REMOVE THIS< COMARES FROM FULL BIVDATA set

# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

#1. determine which strain effects are sig
#use mouse biv average
#Male.poly.Mouse.Table_BivData_4MM

glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm1.SC.male.poly) 
#SC many strains are different (more than the gwRR)
#subsps are sig

glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm2.SC.male.poly)

#Dom
glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(glm_SC.male_Dom)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(glm_SC.male_Musc)

#Mol
glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(glm_SC.male_Mol)

#Logistic regressions

log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))

                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],

#summary(log.SC.male.poly)

#summary(log.SC.male.poly_musc)
#not sig -- when just looking at Musc

#significant in predicting group
log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))

#summary(log.SC.male.poly_mol)
#not sig -- when just looking at Mol
```

@@

```{r SC.length.male_tests, echo=FALSE, include=FALSE}

## run the tests to check following expectations

@@

#Male.poly.Mouse.Table_BivData_4MM
#Long.biv_mouse_table_male.poly

#decide between lm or glm..?
#mouse avs, males

@@

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE}

@@

#pairwise.t.test(Ozone, Month, p.adj = "bonf")

# breief transition, rec landscape

>>>>

total SC t tests
 p = `r Q2tot.sc_t.test_High.Low$p.value`
**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

@@

<within subspecies
musc-short

@@

mol-long
very few significant comparisons>

@@

across subspecies: only significant at the bivalent level, (Dom and Musc are sig different, Dom - Mol are sig difference) (also when high rec and low rec are compared).

@@

```{r Q2.totSC.glm, echo=FALSE}
#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel

@@

@@

glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)
```

@@

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)
#summary(glm_SC.male_Musc_long.biv)

@@

**For the total SC glm**

Thse are our prediction for the glm models based on the variation of the gwRR's

@@

(the subspecies effect is tested in M1)

For total SC in M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)

strain effects from M3; G, LEW, MSM, SKIVE and TOM are significant

@@

@@

The CZECH and AST are higher than KAZ (the first 2 strains drive the lack of difference)

@@

<!--
blocks of other male Q2 - SC lengths
what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.
-->

> **Summary Points**

- confirm that the positive correlation of SC and CO number is maintained

- Predicted differences, between High and low strains OR between Dom and Musc/Mol, only have significant p values when the single bivalent levels are used.

- (the glms, indecate there are many strain effects), Using the mouse average SC lengths, in the full data set all strain effects are significant. This is an indication that there is more variation for the SC lengths than for gwRR / CO counts.

- For the M3 models of mouse av.s across SC meterics, for many of the SC metrics, most all of the strain effects are significant

- for the M2 models of the mouse av, musc - mol are significant fixed effects

- many strain effects, complications in the SC meterics; short/long single bivalents have lower sample size and the total SC might have noise due to the algorithm / differences in the imaging

@@

@@

<These are the plots of the logistic regression.>

@@

@@

@@

```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
## run the tests to check following expectations

@@

@@

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}

@@

#pairwise.t.test(Ozone, Month, p.adj = "bonf")
```

# brief transition, rec landscape
